On The Impact of Atoms of Confusion in JavaScript Code

### 2020-11-11 2_Background.tex
additional review on sections 1, 2

However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions--- when these exist- allowed by the language or framework they are working with. Although this can be, to an extent, a personal matter, since different individuals might hold different opinions on whether a given syntax is difficult to understand, it is possible to empirically investigate if there exist programming languages' constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.

@@

The concept of atoms of confusion was introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. Their work laid an empirical framework to identify such atoms. In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atoms were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.\castor{Mudei um pouco o protocolo para identificar átomos. Tinha achado a explicação anterior confusa.}
%By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.

One main example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a number literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 potential atoms, the authors were able to identify 15 snippets that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study replicates this study, for all atoms that have a counterpart in JavaScript.\castor{Is ``replicates'' the best word here? From what I could grasp, there are some differences between the protocols.} Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on the different nature of confusion amongst the atoms, noting that, for some atoms, a particular incorrect answer was prevalent, whilst there was a distribution of incorrect ones in other atoms.

@@

\castor{Provavelmente devemos reduzir o espaço tomado pelos quatro trabalhos acima, a depender da existência de outros. O de Linguistic Antipatterns talvez até devesse sair.}
\rb{Incluir os trabalhos do grupo do Marcio Ribeiro e do Fernando Castor. Quais outros trabalhos deveriamos incluir aqui?}

>>>>

However, one of the ways in which programs become less comprehensible occurs when programmers favor complicated syntactic constructs instead of using semantically equivalent versions---when these exist---allowed by the language or framework they are working with. Although this can be, to an extent, a personal matter, since different individuals might hold different opinions on whether a given syntax is difficult to understand, it is possible to empirically investigate if there exist programming languages' constructs or idioms that are considerably harder to understand, leading readers of the code to frequently make incorrect predictions of its behavior.

@@

The concept of atoms of confusion was introduced by Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18}, who defined them as small code patterns that can verifiably lead to misunderstandings. In addition, for a code pattern to be considered an atom, there must be some alternative pattern or language construct that is functionally equivalent and less likely to cause confusion. Their work laid an empirical framework to identify such atoms. In summary, the basic approach to determine whether a code pattern is an atom involves: (i) identifying a set of atom candidates; (ii) identifying alternative constructs to these atom candidates; (iii) building a set of small programs that each contain a single atom candidate; (iv) building different versions of these programs where the atoms were removed and which employ the aforementioned alternative constructs; (v) in an experiment or semi-experiment, presenting subjects with the two versions of the programs for each atom candidate and asking subjects to predict their outputs; (vi) statistically comparing the performance of the subjects.
%By having subjects evaluate code where in the expected output relied exclusively on evaluating the result of previously known confusing programs that had been written in \clang or \cpplang, and asking the participants to predict the output of a set of code snippets where the atoms had been removed, the authors were able to isolate and measure the impact of such confusing atoms.

One main example that occurs in many programming languages is the Change of Literal Encoding. For instance, the \clang code statement \lstinline{printf("\%d", 013);} 
often leads programmers to predict the output to be \texttt{13}, even though the correct answer is \texttt{11}. This occurs because a leading \texttt{0} in a number literal indicates that the number is in base \texttt{8}, a fact that is not only unknown to less experienced programmers, but misleading even for seasoned developers. After formulating a list of 19 potential atoms, Gopstein et al. \cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} were able to identify 15 snippets that present a statistically significant difference in comparative answer correctness when each of the 15 atoms was removed. The least confusing atom showed a 14\% boost in prediction accuracy when the atom was removed, whereas the most confusing one showed a 60\% accuracy increase. Our first study follows the empirical framework discussed in~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, for all atoms that have a counterpart in JavaScript. Gopstein et al. \cite{DBLP:conf/msr/GopsteinZFC18} also performed a qualitative analysis on the different nature of confusion amongst the atoms, noting that, for some atoms, a particular incorrect answer was prevalent, whilst there was a distribution of incorrect ones in other atoms.

@@

%% \castor{Provavelmente devemos reduzir o espaço tomado pelos quatro trabalhos acima, a depender da existência de outros. O de Linguistic Antipatterns talvez até devesse sair.}
%% \rb{Incluir os trabalhos do grupo do Marcio Ribeiro e do Fernando Castor. Quais outros trabalhos deveriamos incluir aqui?}
